\section{Potential games}

Potential games form a subclass of non-zero-sum games with guaranteed
pure Nash equilibria (NE) and converging best-response (BR) dynamics,
crucial for decentralized control and exam problems involving
equilibrium existence, computation, and efficiency (e.g., Price of
Anarchy, PoA).

\subsection{N-Player Games and Best-Response Dynamics}

Each player $i$ selects $\gamma_i \in \Gamma_i$
(finite set) to minimize cost:
\[
  J_i( \gamma_1, \gamma_2, ... \gamma_N) = J_i( \gamma_i, \gamma_{-i})
\]

\begin{definition}[Pure Nash equilibrium in $N$-player games]

  A pure strategy profile
  $\gamma^* = (\gamma_1^*, \dots, \gamma_N^*)$
  is a pure NE if:
  \[
    \forall i,\ \forall \gamma_i' \in \Gamma_i: J_i(\gamma_i^*,
    \gamma_{-i}^*) \leq J_i(\gamma_i', \gamma_{-i}^*)
  \]
\end{definition}

\begin{definition}[Pure Best Response in $N$-player games]
  The best (non-empty, possibly non-singleton) response set is:
  \[
    R_i(\gamma_{-i}) = \arg\min_{\gamma_i \in \Gamma_i} J_i(\gamma_i,
    \gamma_{-i})
  \]
  $\gamma^*$ is pure NE iff $\gamma_i^* \in R_i(\gamma_{-i}^*)\ \forall i$
\end{definition}

\textbf{BR dynamics}
Start at $\gamma^0$; iteratively update one player's strategy to a BR,
reducing some $J_i$. May cycle (e.g., rock-paper-scissors)
unless structured (e.g., potential games).

\examBox{Exam relevance}{Prove convergence or find cycles in small games.}

\subsection{Potential Games: Definition and Properties}

\begin{definition}[Exact Potential Function]
  A function $P: \Gamma \to \mathbb{R}$
  is an \textbf{excact potential} if
  $\forall i$, $\forall \gamma_{-i}$,
  $\forall \gamma_i', \gamma_i'' \in \Gamma_i$:
  \[
    J_i(\gamma_i',\gamma_{-i})
    -
    J_i(\gamma_i'',\gamma_{-i})
    =
    P(\gamma_i',\gamma_{-i})
    -
    P(\gamma_i'',\gamma_{-i})
  \]
\end{definition}

A game is a \textbf{potential game} if it admits a potential function.

Key: Unilateral deviation changes match $P$ changes (same for all players).

\begin{proposition}[Existence and Convergence]
  Finite potential games have $\geq 1$ pure NE. BR dynamics converge
  finitely to a pure NE (since $P$ strictly decreases on non-NE
  updates; finite $\Gamma$ prevents infinite descent).
\end{proposition}

Pure NE are local minima of $P$ (directional: no unilateral descent).
Multiple NE possible, initial conditions affect convergence.
\examBox{Exam tip}
{Check potential by verifying cost differences or constructing $P$
(e.g., via summation over paths)}

\subsubsection{Path Improvement Characterization}
A path in $\Gamma$ is a sequence where consecutive profiles differ by
one player's unilateral deviation. A closed path returns to start;
simple if no repeats.

\begin{proposition}
  A game is potential iff the total improvement $I(\mathcal{P}) =
  \sum$ (cost changes along path) is zero for every finite simple
  closed path of length 4 (or all closed paths).
\end{proposition}

\examBox{Proof via cycle decomposition}
{exam-relevant for verifying potential
in small games (e.g., check 4-cycles)}

\subsection{Congestion Games}

Canonical example of potential games, modeling resource sharing
(e.g., traffic, networks).

\begin{definition}[Congestion Game]
  Resources $\{1,\dots,M\}$

  $\gamma_i \subseteq \{1,\dots,M\}$ (subset used by $i$)

  load $\ell_j(\gamma) = |\{i: j \in \gamma_i\}|$

  cost $J_i(\gamma) = \sum_{j \in \gamma_i} f_j(\ell_j(\gamma))$ with
  $f_j$ non-decreasing.
\end{definition}

\begin{theorem}[Rosenthal's Potential]
  Every congestion game is potential with $P(\gamma) = \sum_{j=1}^M
  \sum_{k=1}^{\ell_j(\gamma)} f_j(k)$.
\end{theorem}

\textbf{Proof}:
Unilateral change by $i$ (add/remove resources) alters $P$
exactly by the marginal $f_j(k)$ terms matching $\Delta J_i$.
Thus, pure NE exist; BR converges.

\examBox{Exam example}{Braess paradox in routing
(adding link worsens equilibrium travel time, e.g., from 64 to 70 min)}

\subsection{Equilibrium Refinements and Examples}

Multiple NE: Use Hasse diagram (partial order by Pareto dominance) or
admissibility (no weakly dominated strategies). E.g., in renewable
energy game (moral hazard with fine $f$):

For $c/\epsilon < f < b-s$,
two pure NE ((Reserve, No Shed), (No Reserve, Shed)); one mixed;
Hasse ranks (R,NS) > (NR,S). Security strategies: Maximin (e.g.,
Reserve for player 1).

Identical-interest games (same $J_i = \phi \forall i$) are potential
with $P = \phi$. Zero-sum potential games exist but are trivial or decomposable.

Flight booking: Congestion on tickets; social optimum is NE
(minimizes $P =$ total cost) but PoA $>1$ (counterexample with PoA $\geq 3.5$).

\subsection{Social Welfare and Price of Anarchy}

\begin{definition}[Welfare and PoA]
  Welfare $W(\gamma) = \sum_i J_i(\gamma)$ (or max, etc.). PoA =
  $\max_{\gamma \in \text{NE}} W(\gamma) / \min_{\gamma \in \Gamma}
  W(\gamma)$ (measures inefficiency of worst NE vs optimum).
\end{definition}

In potential games, NE minimize $P$ locally, but $P \neq W$ implies
PoA $>1$ (e.g., Braess: $\approx 1.09$; renewable: PoA=1 with proper
fine aligning incentives).
\examBox{Exam}
{Compute PoA in routing or design mechanisms (fines/lanes) to minimize it}

