\section{Repeated games}

\subsection{Tragedy of the commons}

$N$ players share a single resource.
Each can \textbf{cooperate} (limit consumption)
or \textbf{exploit} (overuse).

For 2 players (maximizers), payoff matrix:
\[
  \begin{matrix}
    & \text{cooperate} & \text{exploit} \\
    \text{cooperate} & (a,a) & (0,b) \\
    \text{exploit} & (b,0) & (c,c) \\
  \end{matrix}
\]

For $N$ players:
\[
  P_i =
  \begin{cases}
    a \frac{N_C - 1}{N-1} & \text{(cooperate, } N_C \text{ cooperators)} \\
    (b-c) \frac{N_C}{N-1} + c & \text{(exploit)}
  \end{cases}
\]
Exploit is dominant strategy and unique pure Nash equilibrium.

\subsubsection{Applications}
Overpopulation, pollution, overfishing, antibiotic resistance, Wi-Fi
overuse, etc.

\subsection{Finitely Repeated Games $G^{(T)}$}

\textbf{Supergame}
$G^{(T)}$, same game $G$ repeated $T$ times.

\textbf{Actions}
$u^t \in \{\text{cooperate},\text{exploit}\}^N$ of $N$ players at stage $t$

\textbf{History of the game}
$\mathcal{H}^t$ actions of $N$ players before stage $t$

$\mathcal{H}^1 = ()$ empty vector,
$\mathcal{H}^T = (u^1,\dots,u^{T-1})$

\textbf{Assumption:}
All agents have perfect recall of the past.

\subsection{Strategies in a repeated game}

\textbf{Pure strategy}
$\gamma_i^{(T)}: \mathcal{H}^t \to \{\text{cooperate, exploit}\}$.

\textbf{Payoff}
$P_i^{(T)} = \sum_{t=1}^T P_i(u^t)$, where $u^t = \gamma^{(t)}(\mathcal{H}^t)$.

NE: No player improves by unilaterally changing strategy (sequence of
functions, not just actions).

\begin{sstTitleBox}{Nash Equilibrium in Finite Repetitions}
  \begin{sstOnlyFrame}
    \begin{theorem}
      If $\gamma^*$ is a NE of $G$, then constant strategy $\gamma^t =
      \gamma^* \ \forall t$ is NE of $G^{(T)}$.
      If $\gamma^*$ is unique NE of $G$, then every NE of $G^{(T)}$
      yields $u^t = \gamma^* \ \forall t$
      (by backward induction: last stage is $G$, induct backwards).
    \end{theorem}
  \end{sstOnlyFrame}
\end{sstTitleBox}

In Tragedy of Commons (unique dominant NE = exploit), only
equilibrium outcome is repeated exploitation; repetition does not
sustain cooperation.

\examBox{Exercise Insight: Repetition Can Lead to Bad Outcomes}
{In some games with unique stage NE (e.g., (A,A) payoff 2 each),
  finite repetition preserves it, but infinite can sustain worse
equilibria (e.g., (B,B) payoff 1 each) via triggers, yielding lower payoffs}

\subsection{Infinitely Repeated Games $G^{(\infty)}$}

\textbf{Discounted payoff}
For patience factor $0 < \delta < 1$:
\[
  P_i^{(\infty)}(\gamma^{(\infty)})
  = (1-\delta) \sum_{t=1}^\infty \delta^{t-1} P_i(u^t)
\]

\textbf{Evil strategy}
$\gamma_i^t = $ exploit $\forall t,i$ is always NE.

\textbf{Grim trigger strategy}\\
$\gamma_i^t = $ cooperate if all cooperated
$\forall \tau < t$, else exploit forever.

\begin{proposition}[Trigger NE in TC]
  For $a=2,b=3,c=1$, all playing grim trigger is NE iff $\delta \geq
  1/2$ (deviation gain $1$ today $\leq$ discounted future loss $1$
  per period: $1 \leq \delta / (1-\delta)$).
\end{proposition}

Tit-for-tat (2 players): Cooperate first, then copy opponent's
previous action; NE for high $\delta$ in PD-like games.

\begin{theorem}[Cooperation via Repetition (Folk Theorem Variant)]
  Let $\hat{\gamma}$ yield $\hat{P}_i > \bar{P}_i :=
  \min_{\gamma_{-i}} \max_{\gamma_i} P_i(\gamma_i, \gamma_{-i})$
  (minmax value) $\forall i$. Then $\exists \delta_0 \in (0,1)$ s.t.
  $\forall \delta > \delta_0$, $G^{(\infty)}$ has NE with on-path
  payoffs $\hat{P}_i$ each period.

  Construction: Grim trigger -- play $\hat{\gamma}$ if no deviations,
  else switch forever to minmax profile against deviator.
\end{theorem}

Infinitely repeated games often have many NE, enabling cooperation if
better than minmax.

\examBox{Exercise Insight: Non-Credible Threats}
{In games with unique stage NE (e.g., (B,A) payoffs (6,0)), infinite
  repetition can sustain better outcomes like repeated (A,A) (4 each)
  via triggers for high $\delta$, but threats may be non-credible
(e.g., punisher hurts themselves more).}

\examBox{Relevance to Exams}
{Focus on: Unique NE preservation in finite games; cooperation
  sustainability in infinite via grim triggers and high $\delta$; folk
  theorem for multiple equilibria; potential pitfalls like bad outcomes
or non-credible threats (hints at subgame perfection).}

